---
title: "512 Project"
author: "Nolan Walker"
date: "2/17/2021"
output: pdf_document
---

```{r setup, include=FALSE,warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***Introduction***

**Data Description**

We are looking at the mouse cortex protein expression dataset from Higuera C et. al(2015), where 77 cortex proteins were observed in a mouse model of down syndrome. Some mice were treated with metamine, a drug used to treat alzheimers, to show that learning can be recovered in down syndrome mice. Mice were either assigned to learn, where they were given context and then a shock, or not to learn, where they were only given a shock. This gives a total of 8 classes(2X2X2).There were 15 measurements for each protein, 38 control mice, and 34 down syndrome mice. The goal of this study is to see if there is a difference in protein expression levels across the different classes for the various proteins.

```{r,eval=TRUE,echo=FALSE,warning=FALSE}
suppressPackageStartupMessages(library(dendextend))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(dendextend))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(tibble))
suppressPackageStartupMessages(library(purrr))
suppressPackageStartupMessages(library(ape))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(broom))
suppressPackageStartupMessages(library(caret))
suppressPackageStartupMessages(library(bnstruct))
suppressPackageStartupMessages(library(naniar))
suppressPackageStartupMessages(library(RANN))
data<-read.csv("C:/Users/Nolan/Documents/Data_Cortex_Nuclear.csv")
```

**Data exploration**

```{r,eval=TRUE}
dim(data)
str(data[-c(2:78)])
unique(data$class)
names(data[c(2:78)])
```
As we can see, there are a lot(1080) of measurements for each of the 77 proteins. There are eight classes; combinations of Genotype, Treatment, and Behavior.

Next to observe missing values.

```{r}
vis_miss(data)
```

As is observed in the plot of missing values, a few of the proteins towards the right of the plot are missing quite a few observations, so some data imputing must be done.


```{r,fig.height=10,fig.width=20,echo=TRUE,eval=TRUE,echo=FALSE}
data.col<-data[,2:78]
data.knn<-preProcess(data.col,method=c("knnImpute"),k=20,knnSummary = mean)
impute_info<-predict(data.knn,data.col,na.action=na.pass)
col.names <- data.frame(col = names(data.knn$mean), mean = data.knn$mean, sd = data.knn$std)
for(i in col.names$col){
 impute_info[i] <- impute_info[i]*data.knn$std[i]+data.knn$mean[i] 
}
Data<-data.frame(impute_info,data$class)
colnames(Data)[colnames(Data)=="data.class"]<-"class"
```

It would be useful to see the distribution of observations for the different proteins.

```{r,fig.height=10,fig.width=20,echo=TRUE,eval=TRUE,echo=FALSE}
par(mfrow=c(3,3))
for (i in 1:77)
  {
   boxplot(data[,i+1]~data$class,xlab="class",ylab="protein levels")
  title(main=names(data)[i+1])
}
```

It looks like some proteins have quite a large difference in measurments across classes. It would make sense to fit multiple mean models to the different proteins to see if there is a difference.

***H_null**: there are no diferrences in expression level mean between classes.*

$$H_0: \mu_0=\mu_1 = \mu_2 = \mu_3 = \mu_4 = \mu_5 = \mu_6 = \mu_7 $$
***H_alt**: there is at least one significant difference among the groups.*

$$H_A: \mu_0 \neq \mu_1 \neq \mu_2 \neq \mu_3 \neq \mu_4 \neq \mu_5 \neq \mu_6 \neq \mu_7 \neq \mu_8$$

**Model fitting**

We fit a simple model to each protein : 

$\mu${Protein|c-CS-m,c-SC-m,c-CS-s,c-SC-s,t-CS-m,t-SC-m,t-CS-s,t-SC-s}$=\beta_0 +\beta_1I_{c-SC-m}+\beta_2I_{c-CS-s}+\beta_3I_{c-SC-s}+\beta_4I_{t-CS-m}+\beta_5I_{t-SC-m}+\beta_6I_{t-CS-s}+\beta_7I_{t-SC-s}$

`lm_herb1`: \[ I_{c-SC-m} = \begin{cases} 
          1 & \text{if class is c-SC-m} \\
          0 & \text{otherwise}
       \end{cases}
      \]
      \[ I_{c-CS-s} = \begin{cases} 
          1 & \text{if class is c-CS-s} \\
          0 & \text{otherwise}
       \end{cases}
      \]
      \[ I_{c-SC-s} = \begin{cases} 
          1 & \text{if class is c-SC-s} \\
          0 & \text{otherwise}
       \end{cases}
      \]
      \[ I_{t-CS-m} = \begin{cases} 
          1 & \text{if class is t-CS-m} \\
          0 & \text{otherwise}
       \end{cases}
      \]
      \[ I_{t-SC-m} = \begin{cases} 
          1 & \text{if class is t-SC-m} \\
          0 & \text{otherwise}
       \end{cases}
      \]
      \[ I_{t-SC-m} = \begin{cases} 
          1 & \text{if class is t-SC-m} \\
          0 & \text{otherwise}
       \end{cases}
      \]
      \[ I_{t-SC-s} = \begin{cases} 
          1 & \text{if class is t-SC-s} \\
          0 & \text{otherwise}
       \end{cases}
      \]
  
```{r,echo=TRUE,eval=TRUE,echo=FALSE,warning=FALSE}
proteins <- names(Data[1:77])
classes <- as.vector(unique(as.character(Data$class))) #creating list of classes names


aov_test <- function(prot){
  return(aov(eval(parse(text = prot)) ~ class, Data) %>% tidy())
}

AOV_df <- data.frame(matrix(NA, nrow = length(proteins), ncol = 3))
names(AOV_df) <- c("Protein", "Var_expl", "p.value")
alpha = 0.01

for(i in 1:length(proteins)){
  #cat("#### ", proteins[i], "\n")
  AOV <- aov_test(proteins[i])
  AOV_df$Protein[i] <- proteins[i]
  AOV_df$Var_expl[i] <- round(AOV$sumsq[1] / (AOV$sumsq[1] + AOV$sumsq[2]), 2)
  AOV_df$p.value[i] <- AOV$p.value
  #cat(paste0("Total variability: ", round(AOV$sumsq[1] + AOV$sumsq[2], 2)))
  #cat('\n\n')
  #cat(paste0("Percentage of explained variability for ", proteins[i], " mean differences: ", round(AOV$sumsq[1] * 100 / (AOV$sumsq[1] + AOV$sumsq[2]), 2), "%"))
  #cat('\n\n')
  #cat(paste0("P value for ", proteins[i], " mean differences: ", round(AOV$p.value[1], 5)))
  #cat('\n\n')
}
AOV_df
```


```{r,fig.height=10,fig.width=20,echo=TRUE,eval=TRUE}
AOV_df$reject_null <- AOV_df$p.value < alpha
alpha_adj <- alpha / length(unique(Data$class))*(length(unique(Data$class))-1)/2
```

