---
title: "STAT 411/511 - Final Stats Report"
author: "Your name"
date: "Date"
output:
  word_document: default
  html_document: default
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) 
library(ggplot2)
```

#### Education versus Income
For the final stats report, we will return to the National Longitudinal Youth Survey dataset.  See Ch 12, Problems 23-24 (and Display 12.20) for a thorough description of the variables included in this dataset.  

From the Sleuth: "In their 1994 book, The Bell Curve: Intelligence and Class Structure in American Life, psychologist Richard Herstein and political scientist Charles Murray arged that a person's intelligence is a better predictor of success in life than is education and family's socioeconomic status.   The book was controversial modely for its conclusions about intelligence and racec, but also for the strength of its conclusions drawn from regression analysis on observational data with imperfect measures of intelligence and socioeconomic statis."

Previously, we used these data to model effects of education level and AFQT score on income in 2005.  Now, I would like you to focus on whether we see an effect of race or gender on income levels, after controlling for education level. To accomplish this, we will combine race and gender variables in the income dataset into a single categorical "Group" variable (See R code below).  


```{r, eval=T, echo=F}
Inc<-Sleuth3::ex1223
  # Response variable
   Inc$Income<-Inc$Income2005 
   # Main explanatory variables
   #Inc$Educ 
   Inc$Group <- paste(Inc$Race, Inc$Gender)
   Inc$Group<-as.factor(Inc$Group)
   levels(Inc$Group)<-factor(c("1F","1M","2F","2M","3F","3M"))
```
```{r}
Inc$Income<-Inc$Income2005 

####### Just a few plots to see if data needed transformin
(p1<-ggplot(Inc,aes(x=Group,y=Income,col=Educ,fill=Educ))+geom_point())
p1+coord_trans(y="log")
p1+coord_trans(y='sqrt')
MASS::boxcox(Income~Group, data=Inc,lambda=seq(.2,.4,length.out = 100))
pairs(Inc[,c('Income','Educ')])

### Income def needs transforming, went with a log
Inc$logInc<-log(Inc$Income)
lm.Int1<-lm(logInc~Group*Educ,data = Inc) ## interaction model
summary(lm.Int1)
lm.add1<-lm(logInc~Group+Educ,data=Inc)### additive model
summary(lm.add1)
confint(lm.add1)
anova(lm.Int1,lm.add1) ## model comparisons, going with additive since they're not sig diff
lm.Male2<-lm(logInc~Educ+I(Group=="1M")+I(Group=="2M")+I(Group=="3M"),data=Inc)
lm.Male<-lm(logInc~Educ+Group,data=subset(Inc,Group==c("1M","2M","3M")))

```
```{r}
par(mfrow=c(2,2))
#plot(lm.Int1)
plot(lm.add1) ## diagnostics

## helps visualize different incomes by group or education
Inc$ED<-factor(ifelse(Inc$Educ < 12, 'L',
                ifelse(Inc$Educ <= 16, 'M', 'H')), levels=c('L','M','H'))
ggplot(Inc,aes(x=Group,y=logInc))+geom_point()+facet_wrap(.~ED,nrow=3)+stat_smooth(method='lm',se=F)
ggplot(Inc,aes(x=Educ,y=logInc))+geom_point()+facet_wrap(.~Group,nrow=3,ncol=3)+stat_smooth(method='lm',se=F)

## nvmnd these 
par(mfrow=c(1,2))
hist(rstandard(lm.add1), freq=F, col='grey', xlim=c(-3,3))
curve(dnorm, xlim=c(-3,3), add=T, lwd=2, col='blue')
plot(lm.add1, which=2) 

par(mfrow=c(1,2))
hist(residuals(lm.Int1), freq=F, col='grey', xlim=c(-8,4))
curve(dnorm, xlim=c(-8,4), add=T, lwd=2, col='blue')
plot(lm.Int1, which=2) 
##

### visualize mean income for different groups
par(mfrow=c(1,1))
boxplot(Inc$logInc~Inc$Group)
```

```{r fig.width=10, fig.height=5}
## income for groups after controlling for education, doesn't show much of a difference
## think this means that after controlling for education the pattern between income and group
## doesn't change much
wch<-c(1,7)
partial.res<-Inc$logInc-model.matrix(lm.add1)[,wch]%*%coef(lm.add1)[wch]
ggpubr::ggarrange(
   ggplot(Inc,aes(x=Group,y=logInc,col=Educ,fill=Educ))+geom_point(),
   
ggplot(data.frame(Inc,partial.res),aes(x=Group,y=partial.res,))+stat_smooth(method='loess', se=F, size=0.5, linetype='dashed')+
   geom_point(),ncol=2)
```
```{r}

```


